---
title: 'Wine datasets EDA'
author: 'lasershow'
date: '10 Jan 2018'
output:
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

## Introduction

This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

[more show](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt)

### ※　Guiding Question

Which chemical properties influence the quality of red wines?


## install package

```{r message = FALSE}
install.packages("ggplot2", dependencies = T)
install.packages("knitr", dependencies = T)
install.packages("dplyr", dependencies = T)
```

```{r}
install.packages('ggthemes')
```

```{r}
install.packages('mice')
```

```{r}
install.packages('randomForest')
```

```{r}
install.packages('tibble')

```

## Load and check data

```{r}
# Load packages
library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
```

Now that our packages are loaded, let's read in and take a peek at the data.

```{r, message=FALSE, warning=FALSE}
data <- read.csv('wineQualityReds.csv', stringsAsFactors = F)

# check data
str(data)
```

## features

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)

2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste

3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines

4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

5 - chlorides: the amount of salt in the wine

6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content

9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

11 - alcohol: the percent alcohol content of the wine

### Output variable (based on sensory data):
12 - quality (score between 0 and 10)

## overview

```{r}
summary(data)
```

## Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
qplot(data=data, x = quality, fill= I("#099DD9")) + 
  ggtitle('Quality')
```

When plotting Quality, 5 and 6 appear most frequently. The summary also shows that the average quality is about 5.6.

### density

Looking at sumarry, it seemed that there was not much difference between maximum and minimum in density. So I plot it and see what kind of distribution it is.

```{r}
qplot(data=data, x = density, fill= I("#099DD9")) + 
  ggtitle('density')
```

Confirming the distribution of density, the average is 0.997. It is understood that it is a normal distribution.

Let's check the binwidth of x in more detail.

```{r}
qplot(data=data, x = density, binwidth = 0.0001, fill= I("#099DD9")) + 
  ggtitle('Density')
```

If I reduce the size of bin, I can see that there is a much count bin nearby 0.99.



