---
title: 'Wine datasets EDA'
date: '10 Jan 2018'
output:
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

## Introduction

This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

[more show](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt)

### ※　Guiding Question

Which chemical properties influence the quality of red wines?

## Load and check data

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load packages
library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
library('corrplot') # visualization
library('GGally')
library('memisc')
```

Now that our packages are loaded, let's read in and take a peek at the data.

```{r echo=FALSE, message=FALSE, warning=FALSE}
data <- read.csv('wineQualityReds.csv', stringsAsFactors = F)

# check data
str(data)
```

## features

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)

2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste

3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines

4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

5 - chlorides: the amount of salt in the wine

6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content

9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

11 - alcohol: the percent alcohol content of the wine

### Output variable (based on sensory data):
12 - quality (score between 0 and 10)

## overview

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(data)
```

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(data=data, x = quality, fill= I("#099DD9")) +
  ggtitle('Quality')
```

When plotting Quality, 5 and 6 appear most frequently. The summary also shows that the average quality is about 5.6.

### density

Looking at sumarry, it seemed that there was not much difference between maximum and minimum in density. So I plot it and see what kind of distribution it is.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(data=data, x = density, fill= I("#099DD9")) +
  ggtitle('density')
```

Confirming the distribution of density, the average is 0.997. It is understood that it is a normal distribution.

Let's check the binwidth of x in more detail.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(data=data, x = density, binwidth = 0.0001, fill= I("#099DD9")) +
  ggtitle('Density')
```

If I reduce the size of bin, I can see that there is a much count bin nearby 0.99.

## volatile.acidit

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(data=data, x=volatile.acidity, xlab="Volatile acidity (g/dm^3)")
```

It is possible to check the deviation value from 1.2 to 1.6.


## chlorides

chlorides is the feature with the biggest difference between the minimum and maximum values. Considering from the average and the median, there is a possibility that it may include outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(data=data, x = chlorides, fill= I("#099DD9")) +
  ggtitle('chlorides')
```

In the box plot, outliers are specified.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(y=data$chlorides, x= 1, geom = "boxplot")
```

Specify the range further.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(y=data$chlorides, x= 1, geom = "boxplot") +
  scale_y_continuous(breaks = seq(0, 0.2, 0.005), limits = c(0, 0.15))

qplot(data=data, x = chlorides,binwidth = 0.001, fill= I("#099DD9")) +
  ggtitle('chlorides')
```

According to the plots, you can see that many values are between 0.06 and 0.09.


## Discussion

I've been searching for one variable so far.
Interestingly, there are few wines with low quality. On the contrary, the high evaluation value of 7 can be seen well.

I also found a variable with outliers. When predicting with machine learning etc, it is necessary to correspond to these outliers.

# Bivariate Plots Section

## correlation

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Correlation Heatmap of Variables
corrplot(cor(data))
```

First, use corrplot to check the correlation. Looking at plot, it turns out that PH has a strong correlation with fixed acidity


## geom density

[geom density document](http://ggplot2.tidyverse.org/reference/geom_density.html)

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Fixed Acidity and Wine Quality
ggplot(data,aes(x=fixed.acidity,fill=factor(quality)))+geom_density(alpha=0.25)+
  scale_x_continuous(breaks = seq(4,16,1))+
  xlab(label = "Fixed Acidity Level")+
  ggtitle("Distribution of Fixed Acidity Levels")+
  theme_few()
```

Based on this plot, in addition to two variables, I can visualize it including quality. However, it is a little hard to see. Therefore, we divide the quality into two categories and visualize it.

### Feature engineering

I will create new features based on quality.
A quality of 6 or more is a good wine, and less than 5 is a bad wine.

[geom bar document](http://ggplot2.tidyverse.org/reference/geom_bar.html)

```{r echo=FALSE, message=FALSE, warning=FALSE}
data$good.wine<-ifelse(data$quality>6,1,0)

#Distribution of good/bad red wines
ggplot(data,aes(x=good.wine,fill=factor(good.wine)))+geom_bar(stat = "count",position = "dodge")+
  scale_x_continuous(breaks = seq(0,1,1))+
  ggtitle("Distribution of Good/Bad Red Wines")+
  theme_few()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Fixed Acidity and Wine Quality
ggplot(data,aes(x=fixed.acidity,fill=factor(good.wine)))+geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(fixed.acidity[good.wine==0],na.rm=T)),color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(fixed.acidity[good.wine==1],na.rm=T)),color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(4,16,1))+
  xlab(label = "Fixed Acidity Level")+
  ggtitle("Distribution of Fixed Acidity Levels")+
  theme_few()
```

Wine with poor quality seems to tend to have higher density when Fixed Acidity Levels is low.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Volatile Acidity and Wine Quality
ggplot(data,aes(x=volatile.acidity,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(volatile.acidity[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(volatile.acidity[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0,1.6,0.1))+
  xlab(label = "Volatile Acidity Level")+
  ggtitle("Distribution of Volatile Acidity Levels")+
  theme_few()
```


When volatile.acidity is low and density is high, it tends to be a bad wine. Also, when volatile.acidity is high, it seems that it tends to be a good wine.
Since a noticeable trend was seen in this plot, let's look at the relation between volatile.acidity and quality in more detail.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(factor(quality), 
            volatile.acidity), 
        data = data) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

It is not that strong, but you can see the correlation with quality.


```{r echo=FALSE, message=FALSE, warning=FALSE}
#Citric Acid and Wine Quality
ggplot(data,aes(x=citric.acid,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(citric.acid[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(citric.acid[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0,1,0.1))+
  xlab(label = "Citric Acid Level")+
  ggtitle("Distribution of Citric Acid Levels")+
  theme_few()
```

If Citric Acid Levels is low and density is high, it tends to be a good wine, in the opposite case it tends to be a bad wine.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Alcohol and Wine Quality
ggplot(data,aes(x=alcohol,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(alcohol[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(alcohol[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(8,15,1))+
  xlab(label = "Alcohol Level")+
  ggtitle("Distribution of Alcohol Levels")+
  theme_few()
```


If alcohol is low and density is high, it tends to be a good wine, in the opposite case it tends to be a bad wine.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(data=data, x=quality, y=total.sulfur.dioxide) +
  geom_point(fill= I('#F79420'), color = I('black'), shape = 21)
```

I plotted the relationship between quality and total.sulfur.dioxide. Outliers can be confirmed when quality is 7.


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = alcohol,
           y = residual.sugar  , color = factor(quality)),
       data = data) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality'))
```

We can see a weak correlation to many quality. In particular, it seems that 3 and 4 have stronger correlation than others.

```{r echo=FALSE, message=FALSE, warning=FALSE}
data$alcohol.bucket = cut(data$alcohol,
                            c(8, 9, 10.5, 12, 16))

ggplot(aes(x = factor(quality),
           y = volatile.acidity ),
           data = data) +
   geom_boxplot( aes(fill= data$alcohol.bucket))  +
  scale_fill_brewer(type='seq',
                   guide=guide_legend(title=data$alcohol.bucket))
```


A negative correlation can be seen from 4 to 7. Interestingly, in 3, volatile.acidity (alcohol 10.5 to 12) is higher than 8, and even 8 (alcohol 12 to 16) is high.
## Discussion

First we checked the correlation with two variables. However, even if two variables have low correlation, we can see that when we look at three variables we have various influences. It is difficult to actually plot four variables, but do not forget that we are in complexity.

# summary

```{r echo=FALSE, Univariate_Plots}
qplot(data=data, x = quality, fill= I("#099DD9")) +
  ggtitle('Quality')
```

As a result of EDA, the distribution of quality of wine was as above. Most of the wine is a quality of 5 or 6, with 7 and 8 being a relatively good wine, 3 and 4 being a bad wine.


```{r echo=FALSE, message=FALSE, warning=FALSE}
#Correlation Heatmap of Variables
data$alcohol.bucket<-NULL
corrplot(cor(data))
```


Looking at the correlation of all the variables, we can see that there is a relatively high correlation between alcohol level and volatile.acidity.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Alcohol and Wine Quality
ggplot(data,aes(x=alcohol,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(alcohol[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(alcohol[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(8,15,1))+
  xlab(label = "Alcohol Level")+
  ggtitle("Distribution of Alcohol Levels")+
  theme_few()
```

In the multivariate plot, we will introduce the difference between the best wine and bad wine remarkably.

This is a diagram visualizing density, alcohol level, quality at once, but when the density and alcohol level are high, you can see that there is a high probability of being a good wine.

## Reflection

The most difficult thing is not only the two variables but also the relationship to quality from the multivariable. The relationship between two variables is comparatively easy to judge by checking the correlation. However, actual things are often more complicated than the relationship of two variables. Therefore, we investigated the relation of three variables this time. According to this survey, it turned out that in the correlation diagram, even variables that were thought to have low correlation values have a relationship when they are three variables.
